\sffamily \Large {\bfseries Introspection dynamics in asymmetric multiplayers games}

### 2023-02-28 latex/article.tex
end of day Tuesday. Need to revise Applications and proof read once more and check marta stuff

%-----SP ABSTRACT-----%
Social interactions are often modelled using simultaneous games. The simplest model involves two identical players. However, interactions in everyday life are complex and often involve more than two individuals that differ between themselves. Such interactions can be captured more realistically using asymmetric multiplayer games. Over the years, learning models and models from evolutionary game theory have studied how players may learn to play strategically in these games. Recently, introspection dynamics has proven to be a useful learning model for tackling complex games. In this dynamics, players compare their current strategy to an alternative strategy. If the update results in a payoff advantage, it is more likely adopted. This dynamics provides a simple way to compute the probability with which players play the different combinations of strategies in the long-run (the stationary distribution). In this paper, we extend some of the previous analytical results of introspection dynamics on games with two-players to games with more players. We obtain analytical expressions of the stationary distribution for multiplayer asymmetric games that have the special property of being additive in the payoffs. In such games, the payoff difference that a player gains by unilaterally switching to a different action is independent of the actions that their co-players play. In addition, we derive the stationary distribution for symmetric multiplayer games with two strategies. We discuss several classical games like the general public goods game to illustrate our results.

@@

\section*{Model of introspection dynamics in multiplayer games}
We consider a normal form game with $N$ players where $N \geq 2$. In the game, a player, say player $i$, can play actions from their action set, $\A_i := \{a_{i,1}, a_{i,2}, ..., a_{i,m_i} \}$. The action set of player $i$ has $m_i$ actions. In this model, players only use pure strategies. Therefore, there are only finitely many states of the game. More precisely, there are exactly $m_1 \times m_2 \times ... \times m_N$ states. We denote a state of the game by collecting the actions of all the players in the game in a vector, $\abf := (a_1, a_2, ..., a_N)$ where $\abf \in \A := \A_1 \times \A_2 \times ... \times \A_N$ and $a_i \in \A_i$. We also use the notation, $\abf := (a_i, \abf_{-i})$ to denote the state from the perspective of player $i$. In the state $(a_i, \abf_{-i})$, player $i$ plays the action $a_i \in \A_i$ and their co-players play the action $\abf_{-i} \in \A_{-i}$ where $\A_{-i}$ is defined as $\A_{-i}:= \prod_{j \neq i} \A_j$. The payoff of a player depends on the state of the game. In this paper, we use bold font letters to denote vectors and matrices. We use the corresponding normal font letters with subscripts to denote elements of the vectors (or matrices). We denote the payoff of player $i$ in the state $\abf$ with $\pi_i(\abf)$ or $\pi_i(a_i, \abf_{-i})$. \\ \\ 
\noindent Since players only use pure strategies in this model, we use the terms strategies and actions interchangebly throughout the whole paper. In this model, players update their strategies over time using the introspection dynamics \cite{Couto:NJP:2022}. At every time step, one randomly chosen player can update their strategy. The randomly chosen player, say $i$, currently playing  action $a_{i,k}$, compares their current payoff to the payoff that they would obtain if they played a randomly selected action,  $a_{i,l} \neq a_{i,k}$, from their action set $\A_i$. This comparison is done while assuming that the co-players do not change their respective actions. When the co-players of player $i$ play $\abf_{-i}$, player $i$ changes from action $a_{i,k}$ to the new action $a_{i,l}$ with the probability, \\
\begin{equation}
 p_{a_{i,k} \to a_{i,l}} (\abf_{-i})= \frac{1}{1 + e^{\displaystyle -\beta(\pi_i(a_{i,l}, \abf_{-i}) - \pi_i(a_{i,k}, \abf_{-i}))}}
 \label{Eq:introspection-update}
\end{equation}
\\ \\ \noindent in the next round. Here $\beta \in [0,\infty)$ is the selection strength parameter that represents the importance that players give to payoff differences while updating their actions. At $\beta = 0$, players update to a randomly chosen strategy with probablity $0.5$. For $\beta > 0$, players update to the alternative strategy under consideration with probablity greater than $0.5$ (or less than $0.5$) if the switch gives them a non-zero increase (or decrease) in the payoffs. The probability that the chosen player does not update the action to a new action is then, \\ \\ 
\noindent Introspection dynamics can be studied by analyzing properties of the transition matrix, $\T$ of the resulting dynamical process. The transition matrix element $\mathrm{T}_{\abf,\bbf}$ denotes the conditional probability that the game goes to the state $\bbf$ in the next round if it is in state $\abf$ in the current round. In order to formally define the transition matrix, we first need to introduce some notations and definitions. We define the neighbourhood set of $\abf$ as the following:

@@

\begin{align}
\mathrm{T}_{\abf, \bbf} = 
\begin{cases}
\frac{1}{N(m_j-1)}  \cdot p_{a_{j} \to b_{j}} (\abf_{-j}) \quad  \quad &\text{ if }\bbf \in \mathrm{Neb}(\abf) \quad \text{and,} \quad j = \mathrm{I}(\abf,\bbf)\\ \\ 
0 \quad &\text{ if } \bbf \notin \mathrm{Neb}(\abf) \\ \\
1 - \sum_{\cbf \neq \bbf} \mathrm{T}_{\abf,\cbf} \quad &\text{ if } \abf = \bbf
\end{cases}
\label{Eq:transition-matrix}
\end{align} \\ 
\noindent The transition matrix is a row stochastic matrix (the sums of the rows are 1). This implies that the stationary distribution of $\T$, a left eigenvector of $\T$ corresponding to eigenvalue $1$, always exists. We introduce a sufficient condition for the stationary distribution of $\T$ to be unique. 
\\ \\
\noindent When the selection strength, $\beta$ is finite, the transition matrix of introspection dynamics has a unique stationary distribution. A finite value of $\beta$ results in non-zero probability of transition between neighbouring states. Since no state is isolated (i.e., every state belongs in the neighbourhood set of another state) and there are only finitely many states of the game, every state is reachable in a finite number of steps from any starting point with non-zero probability. The transition matrix, $\T$, is therefore primitive for a finite $\beta$. By the Perron-Frobenius theorem, a primitive matrix, $\T$ will have a unique and strictly positive stationary distribution $\ubf := (\mathrm{u}_\abf)_{\abf \in \A}$ which satisfies the conditions: 
\begin{eqnarray}
\label{Eq:lefteigenvector}
\ubf \T &= \ubf \\ 
\label{Eq:normalizationcondition}
\ubf \mathbf{1} &= 1
\end{eqnarray}
\noindent where $\mathbf{1}$ is the column vector with size same as $\ubf$ and has all elements as $1$. For all the analytical results in this paper, we consider $\beta$ to be finite so that stationary distribution of the processes are unique. \\ \\
%\begin{Prop} When $\beta$ is finite, the transition matrix of the introspection dynamics has a unique stationary distribution. 
%\label{Prop:unique-stationary-dist}
%\end{Prop}
%\begin{proof}
%A finite value of $\beta$ results in non-zero probability of transition between neighbouring states. Since no state is isolated (i.e., every state belongs in the neighbourhood set of another state) and there are only finitely many states of the game, every state is reachable in a finite number of steps from any starting point with non-zero probability. The transition matrix, $\T$, is therefore primitive for a finite $\beta$. By the Perron-Frobenius theorem, a primitive matrix, $\T$ will have a unique and strictly positive stationary distribution $\ubf := (\mathrm{u}_\abf)_{\abf \in \A}$ which satisfies the conditions: 
%\begin{eqnarray}
%\label{Eq:lefteigenvector}
%\ubf \T &= \ubf \\ 
%\label{Eq:normalizationcondition}
%\ubf \mathbf{1} &= 1
%\end{eqnarray}
%\noindent where $\mathbf{1}$ is the column vector with size same as $\ubf$ and has all elements as $1$. 
%\end{proof}
\noindent The above equations only present an implicit representation of the stationary distribution $\ubf$. The stationary distribution can be explictly calculated by the following expression (which is derived using Eq. (\ref{Eq:lefteigenvector}) and (\ref{Eq:normalizationcondition})) as:
\begin{equation}
\ubf = \mathbf{1}^\intercal (\mathbb{1} + \mathbf{U} - \T)^{-1}
\label{Eq:explicit-stationary-dist-representation}
\end{equation} \\
where $\mathbf{U}$ is a square matrix of the same size as $\T$ with all elements equal to 1 and $\mathbb{1}$ is the identity matrix. The matrix $\mathbb{1} + \mathbf{U} - \T$ is invertible when $\T$ is a primitive matrix \cite{Couto:NJP:2022}. Using Eq. (\ref{Eq:explicit-stationary-dist-representation}) one can compute the stationary distribution of introspection dynamics for any normal form game (with arbitrary number of asymmetric players).\\

\noindent The stationary distribution element $\mathrm{u}_\abf$ is the probability that state $\abf$ will be played by the players in the long run. Using the stationary distribution, one can calculate the marginal probabilities corresponding to each player's actions. For example, the probability that player $i$ plays action $a \in \A_i$ in the long run, $\xi_{i,a}$, can be computed as,
\begin{equation}
\mathbf{\xi}_{i,a} := \sum_{\qbf \in \A_{-i}} \mathrm{u}_{(a, \qbf)}
\label{Eq:marginal-definition}
\end{equation}
%\noindent Since the stationary distribution is a probability distribution, marginal distributions also have the same property. That is, for a player $i$, 
%\begin{equation}
%\sum_{k = 1}^{m_i} \xi_{a_{i,k}}= 1
%\label{Eq:marginal-prob-dist}
%\end{equation}
\section*{Additive games and their properties under introspection dynamics}
In this section we discuss the stationary properties of introspection dynamics when players learn to play strategies in a special class of games: additive games \cite{McAvoy:PlosCB:2015, Pena:JTB:2014}. In an additive game, the payoff difference that a player earns by making a unilateral switch in their actions is independent of what their co-players play. In other words, if none of the co-players change their current actions, the payoff difference earned by making a switch in actions is \emph{only} determined by the switch and not on the actions of the co-players'. Formally, in additive games, for any player $i$, any pair of actions $x,y \in \A_i$, and any $\qbf \in \A_{-i}$,
\begin{equation}
\pi_i(x, \qbf) - \pi_i(y, \qbf) =: f_i(x,y) 
\label{Eq:coplayer-independent-payoffdiff}
\end{equation}
\noindent is independent of $\qbf$ and only dependent on $x$ and $y$. In the literature, this property is sometimes called the property of \emph{equal gains from switching} \cite{Pena:JTB:2014}. For games with this property, the stationary distribution of introspection dynamics takes a simple form;

@@

\label{Th:additive-games-stationary-dist}
\end{Prop}
\noindent For any finite value of the selection strength parameter, $\beta$, the stationary distribution of the process for an additive game can be exactly computed using the expression in Eq. (\ref{Eq:additive-game-stationary-distribution}). For a proof of the propositions and corollaries, please see Appendix. Using the stationary distribution, one can also exactly compute the cumulative probabilities with which players play their actions in the long run (i.e., the marginal distribution) using Eq. (\ref{Eq:marginal-definition}). In this regard, introspection learning in additive games is particularly interesting. The stationary distribution and the marginal distributions of introspection learning in additive games are related in a special way,

@@

\noindent where the marginal probability, $\xi_{j,a_j}$, is the cumulative probability that player $j$ plays $a_j$ at the stationary distribution. The marginal probability for the $N$-player additive game is given by:

\begin{equation}
\xi_{j,a_j} = \frac{1}{\displaystyle \sum_{a' \in \A_j} e^{\beta f_j(a',a_j)}} 
\label{Eq:marginal-at-additive-game}
\end{equation}
\noindent where, $f_j(a', a_j)$ is the co-player independent payoff difference given by Eq. (\ref{Eq:coplayer-independent-payoffdiff}).
\label{Th:additive-game-product-of-marginals}
\end{Prop}
\noindent The above proposition states that for additive games, the stationary distribution of introspection dynamics can be factorized into its corresponding marginals. In the long run, the probability that players play the state $\abf = (a_1, a_2, ...,a_N)$ is the product of the cumulative probabilities that player 1 plays $a_1$, player 2 plays $a_2$ and so on. This property of the additive game was already shown for the simple two-player, two-action donation game in Couto et al. \cite{Couto:NJP:2022}.  Here we extend that result for any number of players, each having an arbitrary number of strategies. In the next section we use the well-studied example of the linear public goods game to demonstrate the stationary properties of introspection dynamics in additive games.

@@

In the simplest version of the linear public goods game with $N-$players, each player has two possible actions, to contribute (action $\C$, to cooperate), or to not contribute (action $\D$, to defect) to the public good. The players differ in their cost of cooperation and the benefit they provide by contributing to the public good. We denote the cost of cooperation for player $i$ and the benefit that they provide to the public good by $c_i$ and $b_i$ respectively. We define an indicator function $\alpha(.)$ to map the action of cooperation to 1 and the action of defection to 0. That is $\alpha(\C) = 1$ and $\alpha(\D) = 0$.  The payoff of player $i$ when the state of the game is $\abf$ is given by: \\

@@

\noindent We use a simple example to illustrate the above result. Consider a 3-player linear public goods game. All players provide a benefit of 2 units when they contribute to the public good ($b_1 = b_2 = b_3 = 2$). They differ, however, in their cost of cooperation. For player 1 and 2, the cost of cooperation is 1 unit ($c_1 = c_2 = 1$) while for the third player, the cost is slightly higher at 1.5 units ($c_3 = 1.5$). In the stationary distribution of the process with selection strength $\beta = 1$, the state where player 1 and 2 play $\C$ and player 3 plays $\D$, $\mathrm{u}_{\C\C\D}$, occurs with probability $\approx 0.121$ (rounded up to three decimal places). We calculate this probability exactly with Eq. (\ref{Eq:stationary_dist_lpgg}). The marginal probabilities that in the long run player 1 (or player 2) plays cooperation and player 3 plays defection are 0.417 and 0.697 respectively. That is, $\xi_{1,\C} = \xi_{2,\C} \approx 0.417$ and $\xi_{3,\D} \approx 0.697$ (rounded up to a three decimal places). With the exact values (and also with the approximations here), one can confirm the factorizing property of the stationary distribution for additive games in this example (i.e., Proposition \ref{Th:additive-game-product-of-marginals}). That is, $\mathrm{u}_{\C\C\D} = \xi_{1,\C} \cdot \xi_{2,\C} \cdot \xi_{3,\D}$. \\ \\
\noindent First, we analyze the simplest case of the linear public goods game in which all players are symmetric to each other. We consider a game with 4 players. All players have the same cost of cooperation ($c_i = c$) and they all provide the same benefit to the public good ($b_i = b$). Since all players are identical, the states of the game can be enumerated by counting the number of cooperators in the state. There can only be 5 distinct states of the game (from 0 to 4 cooperators).
In this case, one can show with Eq. (\ref{Eq:stationary_dist_lpgg}) that, between the states with $k$ and $k+1$ cooperators, the stationary distribution changes by a factor of $(k+1)/(4-k) \cdot \mathrm{exp}(\beta f_j(\D,\C))$. When the parameters of the game are such that defection dominates cooperation ($b = 2, c = 1$, Fig. \ref{Fig:LPGG-symmetric}a), the stationary distribution of the process at high $\beta$ indicates that in the long-run, states with higher number of cooperators are less likely than states with lower number of cooperators. However, for intermediate and low $\beta$, stationary results are qualitatively different. Here, the state with $1$ cooperator (or even $2$ cooperators, depending on how small $\beta$ is) is the most probable state in the long-run (Fig. \ref{Fig:LPGG-symmetric}b).  Naturally, $\beta$ plays an important role in determining the overall cooperation frequncy in the game. When $\beta$ is low, the average cooperation frequency varies weakly with the strength of the dilemma, $b/N - c$ (Fig. \ref{Fig:LPGG-symmetric}c). Even when the temptation to defect is high ($b/N - c \approx -2$), players cooperate with a non-zero probability in the long run ($\approx 0.1$). Similarly, when cooperation is highly beneficial and strictly dominates defection ($b/N - c \approx 2$), players defect with a non-zero probability ($\approx 0.1$). At higher values of $\beta$, the stationary behaviour of players is more responsive to the payoffs and thus reflects an abrupt change around the parameter region where the game transitions from defection-dominating to cooperation-dominating ($b/N - c = 0$).  \\ \\

@@

For studying the asymmetric linear public goods game, we consider a setup with 3 players. All the players can differ in their cost of cooperation and the benefit they provide to the public goods. In this setup, the reference player's (player 2) cost and benefit values are $1$ and $2$ units respectively. Player 1 and player 3 differ from the reference player in opposite directions. For player 1, the cost and benefit are $1 + \delta_c$ and $2 + \delta_b$ respectively while for player 3, the cost and benefit are $1 - \delta_c$ and $2 - \delta_b$, respectively. First, we study a simple case in which players only differ in their cost of cooperation ($\delta_b = 0$ and $\delta_c = 0.5$, Fig \ref{Fig:LPGG-asymmetric}a, left). In the long run, the relative cooperation frequency of players reflect their relative ability to cooperate. The player with the lowest cooperation cost (player 3), cooperates with the highest probability (and \emph{vice-versa},  Fig \ref{Fig:LPGG-asymmetric}a, right). Similarly, when players only differ in their ability to produce the public good ($\delta_b = 1$ and $\delta_c = 0$, Fig \ref{Fig:LPGG-asymmetric}b left), their relative cooperation in the long run reflects the relative benefits they provide with their cooperation. The player with the highest benefit value (player 1) cooperates with the highest probability (Fig \ref{Fig:LPGG-asymmetric}b, right). For this particular setup with three asymmetric players, defection dominates cooperation for player 1 if and only if $\delta_b < 1 + 3\delta_c$ . For player 3, defection dominates cooperation only when $\delta_b > 3\delta_c - 1$. For the parameter values $b = 2, c = 1$, defection always dominates cooperation for the reference player. These regions in the $\delta_b-\delta_c$ parameter plane that correspond to defection dominating cooperation are circumscribed by white dashed lines in Fig. \ref{Fig:LPGG-asymmetric}c. When players learn to play at high selection strength, $\beta$, their cooperation frequency in the long-run reflect the rational play (Fig. \ref{Fig:LPGG-asymmetric}c). In the long run, the average cooperation frequency of the group is low if the asymmetry in the benefit value is bounded, $3\delta_c - 1 < \delta_b < 3\delta_c + 1$. This includes the case where players are symmetric ($\delta_b = \delta_c = 0$). A relatively high cooperation frequency is only assured if players are aligned in their asymmetries (i.e., either $\delta_b < 3\delta_c +1$ or $\delta_b > 3\delta_c - 1$). Or, in other words, if the player that has low cost of cooperation also provides a high benefit upon contribution, then cooperation frequencies are high in the long-run.

>>>>

%-----SP ABSTRACT-----%
Evolutionary game theory and learning models have contributed to the understanding of social interactions and how individuals may learn to act strategically.
The simplest models involve two identical players. However, many interactions in everyday life are often complex, occurring among more than two individuals that differ between themselves. Such interactions can be captured by asymmetric multiplayer games. 
Recently, introspection dynamics has proven to be a useful learning model for tackling asymmetric games. In this dynamics, at each time step, players compare their current strategy to an alternative strategy. If the update results in a payoff advantage, the new strategy is more likely adopted. This model provides a simple way to compute the probability with which players play the different combinations of strategies in the long-run (i.e., the stationary distribution). 
In this paper, we extend some of the previous results of introspection dynamics on 2-player asymmetric games to games with multiple players. First, we derive a general explicit formula that allows us to numerically compute the stationary distribution of any multiplayer asymmetric game. Second, we obtain analytical expressions of the stationary distribution for two special cases, i) additive games (where the payoff difference that a player gains by unilaterally switching to a different action is independent of the actions of their co-players), and ii) symmetric multiplayer games with two strategies. Finally, we revisit several classical games like the general public goods game to illustrate our results.

@@

\section*{Model of introspection dynamics in multiplayer games}
We consider a normal form game with $N$ players where $N \geq 2$. In the game, a player, say player $i$, can play actions from their action set, $\A_i := \{a_{i,1}, a_{i,2}, ..., a_{i,m_i} \}$. The action set of player $i$ has $m_i$ actions. In this model, players only use pure strategies. Therefore, there are only finitely many states of the game. More precisely, there are exactly $m_1 \times m_2 \times ... \times m_N$ states. We denote a state of the game by collecting the actions of all the players in the game in a vector, $\abf := (a_1, a_2, ..., a_N)$ where $\abf \in \A := \A_1 \times \A_2 \times ... \times \A_N$ and $a_i \in \A_i$. We also use the notation, $\abf := (a_i, \abf_{-i})$ to denote the state from the perspective of player $i$. In the state $(a_i, \abf_{-i})$, player $i$ plays the action $a_i \in \A_i$ and their co-players play the action $\abf_{-i} \in \A_{-i}$ where $\A_{-i}$ is defined as $\A_{-i}:= \prod_{j \neq i} \A_j$. In this paper, we use bold font letters to denote vectors and matrices. We use the corresponding normal font letters with subscripts to denote elements of the vectors (or matrices). The payoff of a player depends on the state of the game. We denote the payoff of player $i$ in the state $\abf$ with $\pi_i(\abf)$ or $\pi_i(a_i, \abf_{-i})$. \\ \\ 
\noindent Since players only use pure strategies in this model, we use the terms strategies and actions interchangebly throughout the whole paper. In this model, players update their strategies over time using the introspection dynamics \cite{Couto:NJP:2022}. At every time step, one randomly chosen player can update their strategy. The randomly chosen player, say $i$, currently playing  action $a_{i,k}$, compares their current payoff to the payoff that they would obtain if they played a randomly selected action,  $a_{i,l} \neq a_{i,k}$, from their action set $\A_i$. This comparison is done while assuming that the co-players do not change their respective actions. When the co-players of player $i$ play $\abf_{-i}$, player $i$ changes from action $a_{i,k}$ to the new action $a_{i,l}$ with the probability, \\
\begin{equation}
 p_{a_{i,k} \to a_{i,l}} (\abf_{-i})= \frac{1}{1 + e^{\displaystyle -\beta(\pi_i(a_{i,l}, \abf_{-i}) - \pi_i(a_{i,k}, \abf_{-i}))}}
 \label{Eq:introspection-update}
\end{equation}
\\ \\ \noindent in the next round. Here $\beta \in [0,\infty)$ is the selection strength parameter that represents the importance that players give to payoff differences while updating their actions. At $\beta = 0$, players update to a randomly chosen strategy with probablity $0.5$. For $\beta > 0$, players update to the alternative strategy under consideration with probablity greater than $0.5$ (or less than $0.5$) if the switch gives them a non-zero increase (or decrease) in the payoffs. \\ \\ 
\noindent Introspection dynamics can be studied by analyzing properties of the transition matrix, $\T$ of the resulting dynamical process. The transition matrix element $\mathrm{T}_{\abf,\bbf}$ denotes the conditional probability that the game goes to the state $\bbf$ in the next round if it is in state $\abf$ in the current round. In order to formally define the transition matrix, we first need to introduce some notations and definitions. We start by defining the neighbourhood set of $\abf$ as,

@@

\begin{align}
\mathrm{T}_{\abf, \bbf} = 
\begin{cases}
\frac{1}{N(m_j-1)}  \cdot p_{a_{j} \to b_{j}} (\abf_{-j}) \quad  \quad &\text{ if }\bbf \in \mathrm{Neb}(\abf) \quad \text{and,} \quad j = \mathrm{I}(\abf,\bbf)\\ \\ 
0 \quad &\text{ if } \bbf \notin \mathrm{Neb}(\abf) \\ \\
1 - \sum_{\cbf \neq \bbf} \mathrm{T}_{\abf,\cbf} \quad &\text{ if } \abf = \bbf
\end{cases}
\label{Eq:transition-matrix}
\end{align} \\ 
\noindent The transition matrix is a row stochastic matrix (the sums of the rows are 1). This implies that the stationary distribution of $\T$, a left eigenvector of $\T$ corresponding to eigenvalue $1$, always exists. We introduce a sufficient condition for the stationary distribution of $\T$ to be unique. 
\\ \\
\noindent When the selection strength, $\beta$ is finite, the transition matrix of introspection dynamics has a unique stationary distribution. A finite value of $\beta$ results in non-zero probability of transition between neighbouring states. Since no state is isolated (i.e., every state belongs in the neighbourhood set of another state) and there are only finitely many states of the game, every state is reachable in a finite number of steps from any starting point with non-zero probability. The transition matrix, $\T$, is therefore primitive for a finite $\beta$. By the Perron-Frobenius theorem, a primitive matrix, $\T$ will have a unique and strictly positive stationary distribution $\ubf := (\mathrm{u}_\abf)_{\abf \in \A}$ which satisfies the conditions: 
\begin{eqnarray}
\label{Eq:lefteigenvector}
\ubf \T &= \ubf \\ 
\label{Eq:normalizationcondition}
\ubf \mathbf{1} &= 1
\end{eqnarray}
\noindent where $\mathbf{1}$ is the column vector with size same as $\ubf$ and has all elements as $1$. For all the analytical results in this paper, we consider $\beta$ to be finite so that stationary distributions of the processes are unique. \\ \\
%\begin{Prop} When $\beta$ is finite, the transition matrix of the introspection dynamics has a unique stationary distribution. 
%\label{Prop:unique-stationary-dist}
%\end{Prop}
%\begin{proof}
%A finite value of $\beta$ results in non-zero probability of transition between neighbouring states. Since no state is isolated (i.e., every state belongs in the neighbourhood set of another state) and there are only finitely many states of the game, every state is reachable in a finite number of steps from any starting point with non-zero probability. The transition matrix, $\T$, is therefore primitive for a finite $\beta$. By the Perron-Frobenius theorem, a primitive matrix, $\T$ will have a unique and strictly positive stationary distribution $\ubf := (\mathrm{u}_\abf)_{\abf \in \A}$ which satisfies the conditions: 
%\begin{eqnarray}
%\label{Eq:lefteigenvector}
%\ubf \T &= \ubf \\ 
%\label{Eq:normalizationcondition}
%\ubf \mathbf{1} &= 1
%\end{eqnarray}
%\noindent where $\mathbf{1}$ is the column vector with size same as $\ubf$ and has all elements as $1$. 
%\end{proof}
\noindent The above equations only present an implicit representation of the stationary distribution $\ubf$. The stationary distribution can be explictly calculated by the following expression (which is derived using Eq. (\ref{Eq:lefteigenvector}) and (\ref{Eq:normalizationcondition})) as:
\begin{equation}
\ubf = \mathbf{1}^\intercal (\mathbb{1} + \mathbf{U} - \T)^{-1}
\label{Eq:explicit-stationary-dist-representation}
\end{equation} \\
where $\mathbf{U}$ is a square matrix of the same size as $\T$ with all elements equal to 1 and $\mathbb{1}$ is the identity matrix. The matrix $\mathbb{1} + \mathbf{U} - \T$ is invertible when $\T$ is a primitive matrix \cite{Couto:NJP:2022}. Using Eq. (\ref{Eq:explicit-stationary-dist-representation}) one can compute the unique stationary distribution of (a finite $\beta$-) introspection dynamics for any normal form game (with arbitrary number of asymmetric players).\\

\noindent The stationary distribution element $\mathrm{u}_\abf$ is the probability that state $\abf$ will be played by the players in the long run. Using the stationary distribution, one can calculate the marginal probabilities corresponding to each player's actions. That is, the probability that player $i$ plays action $a \in \A_i$ in the long run, $\xi_{i,a}$, can be computed as,
\begin{equation}
\mathbf{\xi}_{i,a} := \sum_{\qbf \in \A_{-i}} \mathrm{u}_{(a, \qbf)}
\label{Eq:marginal-definition}
\end{equation}
%\noindent Since the stationary distribution is a probability distribution, marginal distributions also have the same property. That is, for a player $i$, 
%\begin{equation}
%\sum_{k = 1}^{m_i} \xi_{a_{i,k}}= 1
%\label{Eq:marginal-prob-dist}
%\end{equation}
\section*{Additive games and their properties under introspection dynamics}
In this section we discuss the stationary properties of introspection dynamics when players learn to play strategies in a special class of games: additive games \cite{McAvoy:PlosCB:2015, Pena:JTB:2014}. In an additive game, the payoff difference that a player earns by making a unilateral switch in their actions is independent of what their co-players play. In other words, if none of the co-players change their current actions, the payoff difference earned by making a switch in actions is \emph{only} determined by the switch and not on the actions of the co-players'. Formally, in additive games, for any player $i$, any pair of actions $x,y \in \A_i$, and any $\qbf \in \A_{-i}$,
\begin{equation}
\pi_i(x, \qbf) - \pi_i(y, \qbf) =: f_i(x,y) 
\label{Eq:coplayer-independent-payoffdiff}
\end{equation}
\noindent is independent of $\qbf$ and only dependent on $x$ and $y$. In the literature, this property is sometimes called the property of \emph{equal gains from switching} \cite{Pena:JTB:2014}. For games with this property, the stationary distribution of introspection dynamics takes a simple form,

@@

\label{Th:additive-games-stationary-dist}
\end{Prop}
\noindent For a proof of the Propositions and Corollaries, please see Appendix. Using the stationary distribution, one can also exactly compute the cumulative probabilities with which players play their actions in the long run (i.e., the marginal distribution) using Eq. (\ref{Eq:marginal-definition}). In this regard, introspection learning in additive games is particularly interesting. The stationary distribution and the marginal distributions of introspection dynamics in additive games are related in a special way,

@@

\noindent where the marginal probability, $\xi_{j,a_j}$, is the cumulative probability that player $j$ plays $a_j$ at the stationary distribution. For the $N$-player additive game, this is given by,

\begin{equation}
\xi_{j,a_j} = \frac{1}{\displaystyle \sum_{a' \in \A_j} e^{\beta f_j(a',a_j)}} 
\label{Eq:marginal-at-additive-game}
\end{equation}
\noindent where, $f_j(a', a_j)$ is the co-player independent payoff difference given by Eq. (\ref{Eq:coplayer-independent-payoffdiff}).
\label{Th:additive-game-product-of-marginals}
\end{Prop}
\noindent The above proposition states that for additive games, the stationary distribution of introspection dynamics can be factorized into its corresponding marginals. In the long run, the probability that players play the state $\abf = (a_1, a_2, ...,a_N)$ is the product of the cumulative probabilities that player 1 plays $a_1$, player 2 plays $a_2$ and so on. This property of the additive game was already shown for the simple two-player, two-action donation game in Couto et al. \cite{Couto:NJP:2022}.  Here we extend that result for any additive game with arbitrary number of players, each having an arbitrary number of strategies. In the next section we use the well-studied example of the linear public goods game (an additive game) to illustrate these results.

@@

In the simplest version of the linear public goods game with $N-$players, each player has two possible actions, to contribute (action $\C$, to cooperate), or to not contribute (action $\D$, to defect) to the public good. The players differ in their cost of cooperation and the benefit they provide by contributing to the public good. We denote the cost of cooperation for player $i$ and the benefit that they provide by $c_i$ and $b_i$ respectively. We define an indicator function $\alpha(.)$ to map the action of cooperation to 1 and the action of defection to 0. That is $\alpha(\C) = 1$ and $\alpha(\D) = 0$.  The payoff of player $i$ when the state of the game is $\abf$ is given by: \\

@@

\noindent We use a simple example to illustrate the above result. Consider a 3-player linear public goods game. All players provide a benefit of 2 units when they contribute to the public good ($b_1 = b_2 = b_3 = 2$). They differ, however, in their cost of cooperation. For player 1 and 2, the cost of cooperation is 1 unit ($c_1 = c_2 = 1$) while for the third player, the cost is slightly higher at 1.5 units ($c_3 = 1.5$). In the stationary distribution of the process with selection strength $\beta = 1$, the cumulative probability that player 1 (or 2) cooperates and player 3 defects are $\xi_{1,\C} = \xi_{2,\C}= 0.417$ and $\xi_{3,\D} = 0.697$ respectively. With the exact values, one can confirm the factorizing property of the stationary distribution for additive games in this example (i.e., Proposition \ref{Th:additive-game-product-of-marginals}). That is, $\mathrm{u}_{\C\C\D} = 0.121 = \xi_{1,\C} \cdot \xi_{2,\C} \cdot \xi_{3,\D}$. \\ \\
\noindent We use the Eq. (\ref{Eq:stationary_dist_lpgg}) to analyze the stationary behaviour of players after a long run of introspection in a linear public goods game (LPGG). First, we study the simplest case where all players are symmetric (the cost and benefit for all the 4 players are $c$ and $b$). Since all players are identical, the states of the game can be enumerated by counting the number of cooperators in the state. There can only be 5 distinct states of the game (from 0 to 4 cooperators).
When the parameters of the game are such that defection dominates cooperation ($b = 2, c = 1$, Fig. \ref{Fig:LPGG-symmetric}a), the stationary distribution of the process at high $\beta$ indicates that in the long-run, states with higher number of cooperators are less likely than states with lower number of cooperators. However, for intermediate and low $\beta$, stationary results are qualitatively different. Here, the state with $1$ cooperator (or even $2$ cooperators, depending on how small $\beta$ is) is the most probable state in the long-run (Fig. \ref{Fig:LPGG-symmetric}b). Since every possible state is equiprobable in the limit of $\beta \to 0$, the outcome with $2$ cooperators is most likely only because there are more states with 2 cooperators than states with any other number of cooperators. \\ \\
\noindent Naturally, $\beta$ plays an important role in determining the overall cooperation in the long run. When $\beta$ is low, average cooperation varies weakly with the strength of the dilemma, $b/N - c$ (Fig. \ref{Fig:LPGG-symmetric}c). Even when the temptation to defect is high ($b/N - c = -2$), players cooperate with a non-zero probability. Similarly, when cooperation is highly beneficial and strictly dominates defection ($b/N - c = 2$), players defect sometimes. At higher values of $\beta$, the stationary behaviour of players is more responsive to the payoffs and thus reflects an abrupt change near the parameters where the game transitions from defection-dominating to cooperation-dominating ($b/N - c = 0$).  \\ \\

@@

To study what effects might appear due to asymmetry in the LPGG, we consider the game with 3 asymmetric players. All the players can differ in their cost of cooperation and the benefit they provide to the public goods. In this setup, the reference player's (player 2) cost and benefit values are $1$ and $2$ units respectively. Player 1 and player 3 differ from the reference player in opposite directions. For player 1, the cost and benefit are $1 + \delta_c$ and $2 + \delta_b$ respectively while for player 3, the cost and benefit are $1 - \delta_c$ and $2 - \delta_b$, respectively. The terms $\delta_b$ and $\delta_c$ represents the strength of asymmetry between the three players (a higher absolute value of $\delta$ indicating a bigger asymmetry). When the players only differ in their cost of cooperation ($\delta_b = 0$ and $\delta_c = 0.5$, Fig \ref{Fig:LPGG-asymmetric}a, left), their relative cooperation in the long run reflects their relative ability to cooperate. The player with the lowest cooperation cost (player 3), cooperates with the highest probability (and \emph{vice-versa},  Fig \ref{Fig:LPGG-asymmetric}a, right). Similarly, when players only differ in their ability to produce the public good ($\delta_b = 1$ and $\delta_c = 0$, Fig \ref{Fig:LPGG-asymmetric}b left), their relative cooperation in the long run reflects the relative benefits they provide with their cooperation (Fig \ref{Fig:LPGG-asymmetric}b, right). In this example, if we consider that the reference player provides a benefit of 2 units and has a cost of 1 unit (in which case, defection always dominates cooperation for them), defection dominates cooperation for player 1 if and only if $\delta_b < 1 + 3\delta_c$ and for player 3 only when $\delta_b > 3\delta_c - 1$. These regions in the $\delta_b-\delta_c$ parameter plane that correspond to defection dominating cooperation are circumscribed by white dashed lines in Fig. \ref{Fig:LPGG-asymmetric}c. When players learn to play at high selection strength, $\beta$, their cooperation frequency in the long-run reflect the rational play (Fig. \ref{Fig:LPGG-asymmetric}c). In the long run, the average cooperation frequency of the group is low if the asymmetry in the benefit value is bounded, $3\delta_c - 1 < \delta_b < 3\delta_c + 1$. This includes the case where players are symmetric ($\delta_b = \delta_c = 0$). A relatively high cooperation is only assured if players are aligned in their asymmetries (i.e., either $\delta_b < 3\delta_c +1$ or $\delta_b > 3\delta_c - 1$). Or, in other words, if the player that has low cost of cooperation also provides a high benefit upon contribution, then cooperation is high in the long-run.
